Double White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-08-13 papers/wdmerger_I/paper.tex
Mike's first round of commments

The first attempts to model the results of the merger process came in the
1980s. \cite{nomotoiben:1985} demonstrated that off-center carbon
ignition would occur in the more massive white dwarf as it accreted
mass near the Eddington rate from the less massive white dwarf
overflowing its Roche lobe. \cite{saionomoto:1985} tracked the
evolution of the flame and found that it propagated quiescently into
the center, converting the carbon-oxygen white dwarf into an
oxygen-neon-magnesium white dwarf. This would then be followed by
collapse into a neutron star---a result with significantly different
observational properties compared to a SN Ia. This scenario, termed
accretion-induced collapse, would be avoided only if the accretion
rate were well below the Eddington rate. \cite{tutukov_yungelson:1979}
observed that this could happen if the mass loss from the secondary
was higher than the Eddington rate and thus the accreted material
formed an accretion disc, which might rain down on the primary more
slowly. The main finding was that double degenerate systems would not
obviously lead to Type Ia supernovae.

@@

The latest developments have resulted in some possible areas of convergence.
\cite{pakmor:2012} performed a merger scenario
with a $1.1\ \msolar$ and $0.9\ \msolar$ setup, with better treatment
of the initial conditions, and indeed found that the merger process
happened over more than ten orbits. Nevertheless, they still determined
that a carbon-oxygen detonation would occur, in line with their
earlier results. \cite{moll:2014} was also able to find a detonation
in a similarly massive system. Very recently, \cite{kashyap:2015} 
found that a detonation occurred in a similarly massive system. Notably,
the detonation occurred self-consistently and did not need to be inserted 
by hand. \cite{dan:2012} and \cite{dan:2014} performed a large sweep 
of the parameter space for merger pairs and
found that pure carbon-oxygen systems would generally not lead to
detonations and violent mergers except for the most massive
systems. They did find that for systems containing helium, many
would detonate and potentially lead to SNe Ia, either through the
aforementioned instabilities in the accretion stream, or during the
contact phase, similar to the violent carbon-oxygen
mergers. \cite{pakmor:2013} added a thin helium shell on their primary
white dwarf, and found that this robustly led to a detonation of the
white dwarf. For now there is preliminary support for the hypothesis
that systems with helium shells (or helium WDs), and very massive carbon-oxygen binaries,
could robustly lead to events resembling SNe Ia.

@@

The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are:
\begin{align}
  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \label{eq:euler_density}\\
  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \label{eq:euler_momentum}\\
  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}). \label{eq:euler_energy}
\end{align}
Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity
vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the
total specific energy, where $e$ is the internal (thermal) specific
energy (energy per unit mass).

@@

A related issue that required a code improvement is that in cases of
large density gradients such as the edge of a star, it is possible to
generate negative densities in zones even with the more strongly
limited PPM. This can occur if a region of large density is moving
away from an ambient zone at relatively large speeds; then the net
density flux in the ambient zones can be large enough to unphysically
drag the density below zero. In practice, this will occur at the
trailing edge of a star that is moving across a grid. In such a
situation, there are two main approaches one could take: either
explicitly introduce a positivity-guaranteeing diffusive flux, or
reset the characteristics of the affected zone. We choose the latter
approach. Even though it is non-conservative, it preserves a
characteristic we value, which is to keep the edge of the stars
relatively sharp, as they physically should be. Since the mass of the
affected zones is typically already fairly low, this should not
seriously affect the energy conservation properties of our
simulation. Our strategy for a reset is as follows: when the density of 
a zone is below a pre-determined density floor (which is typically 
$10^{-5}\ \text{g cm}^{-3}$ for our stellar simulations), we look
at all adjacent zones and find the zone with the highest density.
If it is above the density floor, then we set the field values 
(density, momentum, energy, and temperature) of the
resetted zone to be exactly equal to the field values of this 
adjacent zone. If no adjacent zone reaches the density floor, then
the zone is set to the density floor, and given a temperature equal 
to the temperature floor for our simulations (which is typically 
$10^{5}\ \text{K}$ for our stellar simulations). We then recompute 
the thermodynamics to be consistent with these values. The 
velocity of the zone is set to zero. This latter approach will only
occur in very rare situations, and is mainly there as a safeguard.

\castro's approach to adaptive mesh refinement, based on its underlying
\boxlib\ framework, is to refine zones based on certain user-specified
criteria that tag regions of interest for higher spatial
resolution. Data is represented on one of a number of AMR levels,
where each level corresponds to a set of zones at the same resolution,
which covers a subset of the domain covered by the level immediately
below it. We typically call the level 0 grid the \textit{coarse} grid,
which has the lowest spatial resolution. Each finer, higher-level grid
has a higher resolution than the grid below it by some integer factor
$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in the code. The
zones are strictly contained within the rectangular extent of the
underlying coarser zones (the code is not restricted to representing
only Cartesian geometries, but we use a Cartesian mesh with uniform
spacing in each dimension for the present study). For the time
evolution of the AMR system we use subcycling, where each AMR level is
advanced at a different timestep and a correction step is applied at
the end to synchronize the various levels. Generally the number of
subcycled timesteps is equal to the jump in refinement between levels,
so for example on a grid with three levels and two jumps of four in
refinement, the level 2 zones will have 16 times higher spatial
resolution than the coarse grid and there will be 16 level 2 timesteps
per level 0 timestep.

@@

There are two ways that these source terms enter the hydrodynamics update.
First, during the hydrodynamics update, we alter the edge states that enter
into the determination of the fluxes. (This only applies for the momentum source term;
the gravitational force does not directly do work on the internal energy, which is used 
to infer the pressure.) To second order in space and time, 
this can be done using the cell-centered
source term evaluated at time-level $n$. The details of this are described in 
the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
to the state. First we will describe how we do this for the momentum,
and then we will describe our approach for the energy. This discussion will be somewhat detailed.
However, we believe that the attention is necessary because of the importance of accuracy
in the gravitational source terms for our problem. The stability of the white dwarf binary
system is dependent in large part upon accurate coupling of the hydrodynamics and gravity;
an error in this approach could lead to, for example, a spurious mass transfer episode
that might lead us to very different conclusions about the long term stability of such a system.
Such considerations are generally unimportant for spherically-symmetric single star calculations,
but are of the utmost importance in a simulation where the global gravitational field can change 
quite significantly over the course of the simulation.

@@

In a typical simulation we place the boundaries of the domain far
enough away from the region containing most of the mass that some
method of approximation to this direct summation is justified. Many
approaches exist in the literature. The original release of \castro\
featured the crudest possible approximation: a monopole prescription,
where the boundary values were computed by summing up all the mass on
the domain and treating it as a point source at the domain
center. This is exactly correct only for a spherically symmetric mass
distribution, and therefore is best suited for problems like
single-star Type Ia supernova simulations that employ
self-gravity. However, for a problem like that of a binary star system
with significant departures from spherical symmetry, this assumption
fails to produce accurate boundary values. This results in a
significant drift of the center of the mass of the system over time,
which will be discussed in \autoref{sec:kepler}.

@@

Since the results of a merger simulation depend strongly on gravity,
it is important to check whether proper numerical convergence is
achieved for the Poisson solver. To do so, we created a simple test
that inserts a sphere of radius $R$ and uniform mass density $\rho$
onto our grid, and used \castro\ to calculate the gravitational
potential $\Phi$ of this setup. We ensure that $R$ is an integer
multiple of the grid spacing, and the center of the sphere is at the
origin. The problem domain for our simulations is $[-1.6, 1.6]^3$, and
we take $R = 1.0\ \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
The zones with $r > R$ are filled with an ambient material of very low density 
($10^{-8}\ \text{g cm}^{-3}$). We run this problem at multiple 
resolutions corresponding to jumps by a factor of two. For
comparison, at each grid point we evaluate the analytical potential of
a uniform sphere, which can be easily determined using Gauss' law:
\begin{equation}
  \Phi_{\text{sphere}}(r) = -\frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
\end{equation}
where $M = 4\pi R^3 / 3$ is the mass of the sphere. We measure the 
numerical error by calculating the $L^2$ norm of the error and 
normalizing it by the $L^2$ norm of the analytical solution:
\begin{equation}
  \text{Error} = \frac{\|\Phi - \Phi_{\text{sphere}}\|_2}{\|\Phi_{\text{sphere}}\|_2}.
\end{equation}
We define the order of convergence $p$ between two simulations with a jump 
in resolution of integer factor $m > 1$ as
\begin{equation}
  p = \text{log}_{m}\left(\frac{\text{Error}_{\text{low}}}{\text{Error}_{\text{high}}}\right).
\end{equation}
Here $\text{Error}_{\text{low}}$ is the $L^2$ error at the lower resolution 
and $\text{Error}_{\text{high}}$ is the $L^2$ error at the higher resolution.
We expect the error to converge at $p = 2$ given the discretization we choose. 
For all simulations in this section and for all our main science simulations,
we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{plots/phi_comparison}
  \caption{Comparison of the \castro\ gravitational potential to the analytical solution for: 
    a sphere of uniform density; the same sphere, but with the potential normalized using the 
    actual amount of mass on the grid instead of the mass of a perfect sphere; and, a 
    cube of uniform density. Plotted also is a notional curve whose slope represents
    perfect second order convergence.\label{fig:gravity_convergence}}
\end{figure}
The results of this test are plotted in \autoref{fig:gravity_convergence}.

@@

We can solve one of these two sources of error by changing the density on  
the grid so that the total mass $M$ is actually what we intend it to be.
The resolution study for this case is also
plotted in \autoref{fig:gravity_convergence}. We still obtain
convergence slightly better than second-order, indicating that we 
have not eliminated the geometrical problem of the mass distribution changing.

@@

In a rotating reference frame with angular frequency vector $\bm{\omega}$, the momentum equation becomes:
\begin{equation}
  \left.\frac{\partial(\rho \mathbf{u})}{\partial t}\right|_{\text{rot}} = -2\, {\bm\omega} \times (\rho\mathbf{u}) - \rho {\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right).
\end{equation}
Here $\mathbf{r}$ is the position vector with respect to the origin. Typically we choose $\bm{\omega} = (0, 0, 2\pi / T)^T$,
with the rotation axis coincident with the $z$ axis at $x = y = 0$.
$T$ is the rotation period, which is the most natural quantity to specify
for a rotating stellar system. As described in the appendix, we include this source term
in the edge state prediction in a way that is analogous to the gravity source.
We evaluate all quantities at cell centers. We use the same predictor-corrector 
approach that we use for the gravity source terms to the momentum equations. However, a slight 
difference is that the Coriolis force for each velocity component is coupled to other velocity 
components. If the rotation is about the $z$-axis, then the discrete update to 
$v_x^{n+1}$ depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
would be a slight inconsistency between the $x$ and $y$ components of the velocity.

@@

In all simulations performed in a rotating reference frame, we transform all relevant
quantities back to the inertial reference frame when reporting them. In particular,
for every zone we adjust the momenta and kinetic energy to account for rotation at 
frequency $\omega$. When plotting the locations of 
stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
passed since the simulation started.

@@

@@

We map the 1D model onto the 3D Cartesian grid by taking density,
temperature, and composition as the independent variables,
interpolating these to the cell centers, and then calling the equation
of state to initialize the remaining terms. It is possible to interpolate
instead by using pressure instead of temperature, as pressure is more 
closely related to hydrostatic balance, but the EOS we use is so 
insensitive to temperature that this mapping can result in large 
deviations from the isothermal assumption we started with.  The 
interpolation process divides each zone into $n_{\text{sub}}$ 
sub-zones of equal volume for
the purpose of sampling the 1D model, and the sub-zones are added
together to obtain the full zone's characteristics. This
sub-grid-scale interpolation is useful especially near the edge of the star,
where the density falls off rapidly with radius. Typically we take 
$n_{\text{sub}} = 4$.

>>>>

The first attempts to model the results of the merger process came in the
1980s. \cite{nomotoiben:1985} demonstrated that off-center carbon
ignition would occur in the more massive white dwarf as it accreted
mass near the Eddington rate from the less massive white dwarf
overflowing its Roche lobe.\MarginPar{ref?} \cite{saionomoto:1985} tracked the
evolution of the flame and found that it propagated quiescently into
the center, converting the carbon-oxygen white dwarf into an
oxygen-neon-magnesium white dwarf. This would then be followed by
collapse into a neutron star---a result with significantly different
observational properties compared to a SN Ia. This scenario, termed
accretion-induced collapse, would be avoided only if the accretion
rate were well below the Eddington rate. \cite{tutukov_yungelson:1979}
observed that this could happen if the mass loss from the secondary
was higher than the Eddington rate and thus the accreted material
formed an accretion disc, which might rain down on the primary more
slowly. The main finding was that double degenerate systems would not
obviously lead to Type Ia supernovae.

@@

The latest developments have resulted in some possible areas of convergence.
\cite{pakmor:2012} performed a merger scenario
with a $1.1\ \msolar$ and $0.9\ \msolar$ setup, with better treatment
of the initial conditions, and indeed found that the merger process
happened over more than ten orbits. Nevertheless, they still determined
that a carbon-oxygen detonation would occur, in line with their
earlier results. \cite{moll:2014} was also able to find a detonation
in a similarly massive system. Very recently, \cite{kashyap:2015} 
found that a detonation occurred in a similarly massive system.\MarginPar{``similarly massive system'' repeated twice in succession} Notably,
the detonation occurred self-consistently and did not need to be inserted 
by hand. \cite{dan:2012} and \cite{dan:2014} performed a large sweep 
of the parameter space for merger pairs and
found that pure carbon-oxygen systems would generally not lead to
detonations and violent mergers except for the most massive
systems. They did find that for systems containing helium, many
would detonate and potentially lead to SNe Ia, either through the
aforementioned instabilities in the accretion stream, or during the
contact phase, similar to the violent carbon-oxygen
mergers. \cite{pakmor:2013} added a thin helium shell on their primary
white dwarf, and found that this robustly led to a detonation of the
white dwarf. For now there is preliminary support for the hypothesis
that systems with helium shells (or helium WDs), and very massive carbon-oxygen binaries,
could robustly lead to events resembling SNe Ia.

@@

The Euler equations for hydrodynamics (in the absence of source terms) in conservative form are: \MarginPar{we usually use $U$ for the velocity vector, with $U = (u, v, w)$}
\begin{align}
  \frac{\partial \rho}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}) \label{eq:euler_density}\\
  \frac{\partial \rho \mathbf{u}}{\partial t} &= -\bm{\nabla} \cdot (\rho \mathbf{u}\mathbf{u}) - \bm{\nabla}p \label{eq:euler_momentum}\\
  \frac{\partial \rho E}{\partial t} &= -\bm{\nabla}\cdot(\rho\mathbf{u}E + p\mathbf{u}). \label{eq:euler_energy}
\end{align}
Here $\rho$ is the mass density, $\mathbf{u}$ is the fluid velocity
vector, $p$ is the pressure, and $E = \mathbf{u}^2 / 2 + e$ is the
total specific energy, where $e$ is the internal (thermal) specific
energy (energy per unit mass).

@@

A related issue that required a code improvement is that in cases of
large density gradients such as the edge of a star, it is possible to
generate negative densities in zones even with the more strongly
limited PPM. This can occur if a region of large density is moving
away from an ambient zone at relatively large speeds; then the net
density flux in the ambient zones can be large enough to unphysically
drag the density below zero. In practice, this will occur at the
trailing edge of a star that is moving across a grid. In such a
situation, there are two main approaches one could take: either
explicitly introduce a positivity-guaranteeing diffusive flux, or
reset the characteristics \MarginPar{``characteristics'' has a very precise meaning for hyperbolic systems, use a different word} of the affected zone. We choose the latter
approach. Even though it is non-conservative, it preserves a
characteristic we value, which is to keep the edge of the stars
relatively sharp, as they physically should be. Since the mass of the
affected zones is typically already fairly low, this should not
seriously affect the energy conservation properties of our
simulation. Our strategy for a reset is as follows: when the density of 
a zone is below a pre-determined density floor (which is typically 
$10^{-5}\ \text{g cm}^{-3}$ for our stellar simulations), we look
at all adjacent zones and find the zone with the highest density.
If it is above the density floor, then we set the field values 
(density, momentum, energy, and temperature) of the
reset zone to be exactly equal to the field values of this 
adjacent zone. If no adjacent zone reaches the density floor, then
the zone is set to the density floor, and given a temperature equal 
to the temperature floor for our simulations (which is typically 
$10^{5}\ \text{K}$ \MarginPar{do we really need it this cold?  I thought we were doing $10^6$~K} for our stellar simulations). We then recompute 
the thermodynamics to be consistent with these values. The 
velocity of the zone is set to zero. This latter approach will only
occur in very rare situations, and is mainly there as a safeguard.

\castro's approach to adaptive mesh refinement, based on its underlying
\boxlib\ framework, is to refine zones based on certain user-specified
criteria that tag regions of interest for higher spatial
resolution. Data is represented on one of a number of AMR levels,
where each level corresponds to a set of zones at the same resolution,
which covers a subset of the domain covered by the level immediately
below it. We typically call the level 0 grid the \textit{coarse} grid,
which has the lowest spatial resolution. Each finer, higher-level grid
has a higher resolution than the grid below it by some integer factor
$N$, which is restricted to be $N = 2\ \text{or}\ 4$ in \castro. The
zones are strictly contained within the rectangular extent of the
underlying coarser zones (the code is not restricted to representing
only Cartesian geometries, but we use a Cartesian mesh with uniform
spacing in each dimension for the present study). For the time
evolution of the AMR system we use subcycling, where each AMR level is
advanced at a different timestep and a correction step is applied at
the end to synchronize the various levels. Generally \MarginPar{why ``Generally''?  Is there ever as case where the subcycled step isn't the jump in refinement?} the number of
subcycled timesteps is equal to the jump in refinement between levels,
so for example on a grid with three levels and two jumps of four in
refinement, the level 2 zones will have 16 times higher spatial
resolution than the coarse grid and there will be 16 level 2 timesteps
per level 0 timestep.

@@

There are two ways that these source terms enter the hydrodynamics update. \MarginPar{``hydrodynamics update'' appears twice in succession}
First, during the hydrodynamics update, we alter the edge states that enter
into the determination of the fluxes. (This only applies for the momentum source term;
the gravitational force does not directly do work on the internal energy, which is used 
to infer the pressure.) To second order in space and time, 
this can be done using the cell-centered
source term evaluated at time-level $n$. \MarginPar{we trace, so it's not just the cell-centered source term} The details of this are described in 
the appendix. Second, after the hydrodynamics step, we add the time-centered source terms
to the state. First we will describe how we do this for the momentum,
and then we will describe our approach for the energy. This discussion will be somewhat detailed.
However, we believe that the attention is necessary because of the importance of accuracy
in the gravitational source terms for our problem. The stability of the white dwarf binary
system is dependent in large part upon accurate coupling of the hydrodynamics and gravity;
an error in this approach could lead to, for example, a spurious mass transfer episode
that might lead us to very different conclusions about the long term stability of such a system.
Such considerations are generally unimportant for spherically-symmetric single star calculations,
but are of the utmost importance in a simulation where the global gravitational field can change 
quite significantly over the course of the simulation.

@@

In a typical simulation we place the boundaries of the domain far
enough away from the region containing most of the mass that some
method of approximation to this direct summation is justified. Many
approaches exist in the literature. The original release of \castro\
featured the crudest possible approximation: a monopole prescription,
where the boundary values were computed by summing up all the mass on
the domain and treating it as a point source at the domain
center. This is exactly correct only for a spherically symmetric mass
distribution, and therefore is best suited for problems like
single-star Type Ia supernova simulations that employ
self-gravity. \MarginPar{also try for Castro core-collapse sims---maybe reference a paper on each subject from castro} However, for a problem like that of a binary star system
with significant departures from spherical symmetry, this assumption
fails to produce accurate boundary values. This results in a
significant drift of the center of the mass of the system over time,
which will be discussed in \autoref{sec:kepler}.

@@

Since the results of a merger simulation depend strongly on gravity,
it is important to check whether proper numerical convergence is
achieved for the Poisson solver. To do so, we created a simple test
that inserts a sphere of radius $R$ and uniform mass density $\rho$
onto our grid, and used \castro\ to calculate the gravitational
potential $\Phi$ of this setup. We ensure that $R$ is an integer
multiple of the grid spacing, and the center of the sphere is at the
origin. The problem domain for our simulations is $[-1.6 \text{cm}, 1.6 \text{cm}]^3$, and
we take $R = 1.0\ \text{cm}$ and $\rho = 10^3\ \text{g cm}^{-3}$. 
The zones with $r > R$ are filled with an ambient material of very low density 
($10^{-8}\ \text{g cm}^{-3}$). We run this problem at multiple 
resolutions corresponding to jumps by a factor of two. For
comparison, at each grid point we evaluate the analytical potential of
a uniform sphere, which can be easily determined using Gauss' law:
\begin{equation}
  \Phi_{\text{sphere}}(r) = -\frac{GM}{r} \times \begin{cases} (3R^2 - r^2)/(2 r^2) & r \leq R \\ 1 & r > R \end{cases},\label{eq:sphere-analytical}
\end{equation}
where $M = 4\pi R^3 / 3$ is the mass of the sphere. We measure the 
numerical error by calculating the $L^2$ norm of the error and 
normalizing it by the $L^2$ norm of the analytical solution:
\begin{equation}
  \text{Error} = \frac{\|\Phi - \Phi_{\text{sphere}}\|_2}{\|\Phi_{\text{sphere}}\|_2}.
\end{equation}
We define the order of convergence $p$ between two simulations with a jump 
in resolution of integer factor $m > 1$ as
\begin{equation}
  p = \text{log}_{m}\left(\frac{\text{Error}_{\text{low}}}{\text{Error}_{\text{high}}}\right).
\end{equation}
Here $\text{Error}_{\text{low}}$ is the $L^2$ error at the lower resolution 
and $\text{Error}_{\text{high}}$ is the $L^2$ error at the higher resolution.
We expect the error to converge at $p = 2$ given the discretization we choose. 
For all simulations in this section and for all our main science simulations,
we choose a relative error tolerance of $10^{-10}$ to be satisfied in the multigrid solve.
\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{plots/phi_comparison}
  \caption{Comparison of the \castro\ gravitational potential to the analytical solution for: 
    a sphere of uniform density; the same sphere, but with the potential normalized using the 
    actual amount of mass on the grid instead of the mass of a perfect sphere; and, a 
    cube of uniform density. Plotted also is a notional curve whose slope represents
    perfect second order convergence.\label{fig:gravity_convergence}}
\end{figure}
The results of this test are plotted in \autoref{fig:gravity_convergence}.

@@

We can solve one of these two sources of error by changing the density on  
the grid so that the total mass $M$ is actually what we intend it to be.
The resolution study for this case is also
plotted in \autoref{fig:gravity_convergence}. \MarginPar{explain that this is the ``normalized sphere'' in the figure} We still obtain
convergence slightly better than second-order, indicating that we 
have not eliminated the geometrical problem of the mass distribution changing.

@@

In a rotating reference frame with angular frequency vector $\bm{\omega}$, the momentum equation becomes:
\begin{equation}
  \left.\frac{\partial(\rho \mathbf{u})}{\partial t}\right|_{\text{rot}} = -2\, {\bm\omega} \times (\rho\mathbf{u}) - \rho {\bm\omega} \times \left({\bm\omega} \times \mathbf{r}\right).
\end{equation}
Here $\mathbf{r}$ is the position vector with respect to the origin. Typically we choose $\bm{\omega} = (0, 0, 2\pi / T)^T$,
with the rotation axis coincident with the $z$ axis at $x = y = 0$.
$T$ is the rotation period, which is the most natural quantity to specify
for a rotating stellar system. As described in the appendix, we include this source term
in the edge state prediction in a way that is analogous to the gravity source.
We evaluate all quantities at cell centers. We use the same predictor-corrector 
approach that we use for the gravity source terms to the momentum equations. However, a slight 
difference is that the Coriolis force for each velocity component is coupled to other velocity 
components. If the rotation is about the $z$-axis, then the discrete update to 
$v_x^{n+1}$ \MarginPar{we are using $u, v, w$ below, and is more consistent with other papers / User's Guide} depends on the value of $v_y^{n+1}$, and vice versa. If we fix the value of 
the time-level $n+1$ quantities after coming out of the hydrodynamics update, there 
would be a slight inconsistency between the $x$ and $y$ components of the velocity.

@@

In all simulations performed in a rotating reference frame, we transform all relevant
quantities back to the inertial reference frame when reporting them. In particular,
for every zone we adjust the momenta and kinetic energy to account for rotation at 
frequency $\omega$. When plotting the locations of 
stars on the grid, we rotate them by an angle corresponding to the amount of time that has 
passed since the simulation started. \MarginPar{show the transforms.  Also, be more explicit---is this just done for the plotfiles?}

@@

\MarginPar{use subsections to break the distinct topics apart in this section---it all runs together}

@@

We map the 1D model onto the 3D Cartesian grid by taking density,
temperature, and composition as the independent variables,
interpolating these to the cell centers, and then calling the equation
of state to initialize the remaining terms. It is possible to interpolate
instead by using pressure instead of temperature, as pressure is more 
closely related to hydrostatic balance, but the EOS we use is so 
insensitive to temperature that this mapping can result in large 
deviations from the isothermal assumption we started with.  The 
interpolation process divides each zone into $n_{\text{sub}}$ 
sub-zones of equal volume for
the purpose of sampling the 1D model, and the sub-zones are added
together to obtain the full zone's characteristics \MarginPar{use ``state'' instead of ``characteristics''}. This
sub-grid-scale interpolation is useful especially near the edge of the star,
where the density falls off rapidly with radius. Typically we take 
$n_{\text{sub}} = 4$.
